Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/design.tex
More examples

Nail's fundamental parsers represent signed or unsigned
integers with arbitrary lengths up to 64 bits, an
approach popularized by the Hammer framework~\cite{hammer-parser}.
Note that is possible to define parsers for sub-byte lengths, e.g.,
to parse the 4-bit
data offset within the TCP header.  The internal representation rounds up to the nearest integer
type; for example, a 24-bit unsigned integer will be stored in an unsigned 32-bit integer.

@@

Nail provides a structure combinator with semantic labels instead of the sequence combinator that
other parser combinator libraries use to capture structures in data formats. 
The structure combinator consists of a sequence of fields, typically consisting of a label and a
parser that describes the contents of that field. Other field types will be described below.
 The syntax of the structure combinator is inspired by the Go language~\cite{golang}, with field names preceding their definition.
In our opinion, this syntax is easier to read than the widespread C-style structure and type
declarations.

@@

\paragraph{Optional.}
Nail includes an \cc{optional} combinator, which attempts to recognize a value, but succeeds
without consuming input when it cannot recognize that value. Syntactically, \cc{optional} is
equivalent to a choice between the parser and an empty structure, but in the internal model it is
more concisely represented as a reference that is null when the parser fails.

\paragraph{Rules.}
A Nail grammar consists of rules that assign a parser to a name. Rules are written as
assignments, such as \cc{ints = many uint8}, which defines a rule called \cc{ints} with the
body \cc{many uint8}.  As we will describe in \S\ref{s:dependent}
and \S\ref{s:transforms}, rules can optionally consume parameters. 
Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
the body of the rule had been substituted in the code. If parameters appear, they are passed by reference.

\paragraph{References.}
Rules allow for recursive grammars. To support recursive data types, we allow introduce the
reference combinator \cc{*}  that does not change the syntax of the external format described, but
introduces a layer of indirection, such as a reference or pointer, to the model data type.
For example example,   balanced parentheses could be parsed as follows: \cc{paren = < uint8="("; optional *paren; uint8=")" >}.

@@

The programmer has to guarantee memory safety and unambiguity of the stream transformations. To
reduce both programmer effort and the risk of unsafe operations, we provide implementations of
transformations for many common features, such as checksums, size, and offset fields. Furthermore,
we provide library functions that can be used to safely operate on streams, such as splitting and
concatenation. Nail implements streams as iterators, so they can share underlying buffers and can be
efficiently duplicated and split.

Transformations can handle a wide variety of patterns in data formats, including the following: 
\paragraph{Offsets.}
We provide a transformation for handling offset fields, which is invoked as follows:
\cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 
This transformation corresponds to two functions for parsing and
generation, as shown in Figure~\ref{fig:xform-sig}. It defines a new stream \cc{\$fragment} that can
be used to parse data at the offset contained in \cc{@offset}, by
using \cc{apply \$fragment some_parser}.

@@

\paragraph{Compressed data.}
Encoded, compressed, or encrypted data can be handled transparently by writing a custom
transformation that transforms a coded stream into one that can be parsed by a Nail grammar and vice
versa. Usually, this involves calling an existing library.

>>>>

Nail's fundamental parsers represent signed or unsigned
integers with arbitrary lengths up to 64 bits.
Note that is possible to define parsers for sub-byte lengths, e.g.,
to parse the 4-bit
data offset within the TCP header.  The internal representation rounds up to the nearest integer
type; for example, a 24-bit unsigned integer will be stored in an unsigned 32-bit integer.

@@

Nail provides a structure combinator with semantic labels instead of the sequence combinator that
other parser combinator libraries use to capture structures in data formats. 
The structure combinator consists of a sequence of fields, typically consisting of a label and a
parser that describes the contents of that field. Other field types will be described below.
 The syntax of the structure combinator is inspired by the Go language~\cite{golang}, with field names preceding their definition.

@@

\paragraph{Optional.}
Nail includes an \cc{optional} combinator, which attempts to recognize a value, but succeeds
without consuming input when it cannot recognize that value. Syntactically, \cc{optional} is
equivalent to a choice between the parser and an empty structure, but in the internal model it is
more concisely represented as a reference that is null when the parser fails.
For example, the grammar for  Ethernet headers uses \cc{optional vlan\_header} to parse the VLAN
header that only appears in Ethernet packets transmitted to a non-default VLAN.
\paragraph{Rules.}
A Nail grammar consists of rules that assign a parser to a name. Rules are written as
assignments, such as \cc{ints = many uint8}, which defines a rule called \cc{ints} with the
body \cc{many uint8}.  As we will describe in \S\ref{s:dependent}
and \S\ref{s:transforms}, rules can optionally consume parameters. 
Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
the body of the rule had been substituted in the code. If parameters appear, they are passed by reference.

\paragraph{References.}
Rules allow for recursive grammars. To support recursive data types, we allow introduce the
reference combinator \cc{*}  that does not change the syntax of the external format described, but
introduces a layer of indirection, such as a reference or pointer, to the model data type.
The reference combinator does not need to used when another combinator, such as \cc{optional} or
\cc{many} already introduces indirection in the data type.
For example example,   arithmetic expressions consisting of sums and products can be parsed as follows:

\begin{verbatim}
expr = choose {
  PAREN = <uint8='('; *expr; uint8=')'>
  PRODUCT = sepBy1 uint8='*' expr
  SUM = sepBy1 uint8='+' expr
  INTEGER = many1 uint8 | '0' .. '9'
}
\end{verbatim} \cc{paren = < uint8="("; optional *paren; uint8=")" >}.

@@

 To reduce both programmer effort and the risk of unsafe operations, Nails provides implementations of
transformations for many common features, such as checksums, size, and offset fields. Furthermore,
Nail provides library functions that can be used to safely operate on streams, such as splitting and
concatenation. Nail implements streams as iterators, so they can share underlying buffers and can be
efficiently duplicated and split.

Transformations need to be carefully written, because they can escape Nail's security and introduce
bugs. However, as we will show in \S\ref{s:eval-effort}, Nail transformations are much shorter than
hand-written parsers and many formats can be represented with just the provided transforms

Transformations can handle a wide variety of patterns in data formats, including the following: 
\paragraph{Offsets.}
A built-in transformation for handling offset fields, which is invoked as follows:
\cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 
This transformation corresponds to two functions for parsing and
generation, as shown in Figure~\ref{fig:xform-sig}. It defines a new stream \cc{\$fragment} that can
be used to parse data at the offset contained in \cc{@offset}, by
using \cc{apply \$fragment some_parser}.

@@

\paragraph{Compressed data.}
Encoded, compressed, or encrypted data can be handled transparently by writing a custom
transformation that transforms a coded stream into one that can be parsed by a Nail grammar and vice
versa. This transformation must be carefully written to not have bugs.

### 2014-05-01 paper/osdi14/intro.tex
More examples

First, parser generators typically parse inputs into an abstract syntax
tree (AST) that corresponds to the grammar.  In order to produce a data
structure that the rest of the application code can easily process,
application developers must write explicit {\em semantic actions} that
update the application's internal representation of the data based on
each AST node.  Writing these semantic actions by hand is error-prone,
much like other input processing code, and mistakes can result in memory
corruption bugs or misinterpreted inputs.
\XXX[do we have any evidence to cite for this claim?]
Writing these semantic actions
also requires the programmer to describe the structure of the input three
times---once to describe the grammar, once to describe the internal data
structure, and once again in the semantic actions that translate the
grammar into the data structure---leading to another potential source
of bugs and inconsistencies.

@@

To evaluate whether Nail's design is effective at handling real-world
data formats, we implemented a prototype of Nail for C\@.  Using our
prototype, we implemented succinct grammars for parts of an IP network
stack, for DNS packets, and for ZIP files, each in about a hundred lines
of grammar.  On top of these grammars, we were able to build a fully
functional DNS server in under 200 lines of C code, and an {\tt unzip}
utility in about 50 lines of C code.  This suggests both that Nail is
effective at handling complex real-world data formats, and that Nail
makes it easy for application developers to parse and generate external
data representations.  Performance results show that the Nail-based DNS
server outperforms the widely used BIND DNS server, demonstrating that
Nail-based parsers and generators can achieve good performance.

>>>>

First, parser generators typically parse inputs into an abstract syntax
tree (AST) that corresponds to the grammar.  In order to produce a data
structure that the rest of the application code can easily process,
application developers must write explicit {\em semantic actions} that
update the application's internal representation of the data based on
each AST node.  Not only does the programmer allocate memory and manipulate pointers,
 operations that are error-prone, but
writing these semantic actions
also requires the programmer to describe the structure of the input three
times---once to describe the grammar, once to describe the internal data
structure, and once again in the semantic actions that translate the
grammar into the data structure---leading to another potential source
of bugs and inconsistencies.

@@

To evaluate whether Nail's design is effective at handling real-world data
formats, we implemented a prototype of Nail for C\@. Using our prototype, we
implemented succinct grammars for parts of an IP network stack, for DNS packets,
and for ZIP files, each in about a hundred lines of grammar. On top of these
grammars, we were able to build a DNS server in under 200 lines of C code, and
an {\tt unzip} utility in about 50 lines of C code, with performance comparable
to or exceeding existing implementations.This suggests both that Nail is
effective at handling complex real-world data formats, and that Nail makes it
easy for application developers to parse and generate external data
representations. Performance results show that the Nail-based DNS server
outperforms the widely used BIND DNS server, demonstrating that Nail-based
parsers and generators can achieve good performance.

### 2014-05-01 paper/osdi14/relwk.tex
More examples

Hammer implements grammars as language-integrated parser combinators, an approach popularized by
Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our knowledge, first
described by Burge~\cite{burge1975recursive}) is a natural way of concisely expressing top-down
grammars~\cite{Danielsson:2010:TPC:1863543.1863585}
by composing them from one or multiple sub-parsers.\footnote{For more
background on the history of expressing grammars, see Bryan Ford's
masters thesis~\cite{ford2002packrat}, which also describes the
default parsing algorithm used by Hammer.}
Hammer then constructs a tree of function pointers which can be invoked to parse a given input into
an AST\@.

>>>>

Hammer implements grammars as language-integrated parser combinators, an approach popularized by
Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our knowledge, first
described by Burge~\cite{burge1975recursive}) is a natural way of concisely e               xpressing top-down
grammars~\cite{Danielsson:2010:TPC:1863543.1863585}
by composing them from one or multiple sub-parsers.\footnote{For more
background on the history of expressing grammars, see Bryan Ford's
masters thesis~\cite{ford2002packrat}, which also describes the
default parsing algorithm used by Hammer.}
Hammer then constructs a tree of function pointers which can be invoked to parse a given input into
an AST\@.
